\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{url}

%SetFonts

%SetFonts


\title{Homework 2}
\author{CS 4364/5364\\Spring 2021}
\date{Due: 25 February 2021}							% Activate to display a given date or no date

\begin{document}
\maketitle

Because of the reliance of the particular assignments in this class on mathematical notation, 
and the fact that all assignments will be submitted electronically, 
students are encouraged to use \LaTeX{} to formalize their responses. 
\textbf{For those enrolled in the graduate section the use of latex is \emph{required}.}
This assignment (like all others) will be posted on the course \texttt{github}\footnote{\url{github.com/deblasiolab/CS4364-documents}} as source code as well as in PDF form on the course website. 
Please submit your assignment to the professor via email, either as a link to your assignment online (i.e. overleaf or github) or as an attachment. 
Graduate students will need to include the \texttt{.tex} files as well as a PDF, this is optional but encouraged for undergraduates. 


\begin{enumerate}


\item \textbf{(20 points)} 
Given two sequences $S$ and $T$ (not necessarily the same length), 
let $G$, $L$, and $H$ be the scores of an optimal global alignment, an optimal local alignment, 
and an optimal global alignment without counting the indels at the beginning of $S$ and the end of $T$, respectively. 
\begin{enumerate}
\item Give an example of $S$ and $T$ so that all 3 scores, $G$, $L$, and $H$, are different. 
\item Prove or disprove the statement $L \ge H \ge G$.
\end{enumerate}

Note: for the proof, this can be a sketch, you just need to convince me that you're correct. 

%\item \textbf{(1 Point)}
%Under the affine gap model, describe an algorithm that find the optimal \textit{local} alignment of strings $S[1...n]$ and $T[1...m]$ in $O(mn)$-time. 

\item \textbf{(20 points)} 
Given two strings $S[1...n]$ and $T[1...m]$, we would like to find the two non-overlapping alignments 
($S[i_1...i_2],T[j_1...j_2]$) and ($S[i_3...i_4],T[j_3...j_4]$) such that $i_2<i_3$ and $j_2<j_3$ and the total alignment score is maximized 
in running time $O(mn)$. 
Hints: remember that the local alignment between two sequences is the alignment of a pair of substrings from $S$ and $T$ such that the alignment score is maximized, 
and that the question does not say anything about the relationship between $i_1$ and $i_2$ nor $i_3$ and $i_4$ that if $i_3>i_4, S[i_3...i_4]$ is the empty string.
\end{enumerate}


\end{document}  