\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage[margin=1in]{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{tcolorbox}
\usepackage{url}
\usepackage[normalem]{ulem}
\usepackage{xcolor}
\usepackage{amsmath}
%SetFonts
%\usepackage{multicolumn}
\usepackage{multirow}
\usepackage{xspace}
%SetFonts
\newcommand{\change}[2]{\sout{#1}\xspace\textcolor{orange}{#2}}
\newcommand{\add}[1]{\textcolor{orange}{#1}}


\title{Homework 9}
\author{CS 4364/5364\\Spring 2022}
\date{Due: 9 May 2022}							% Activate to display a given date or no date

\begin{document}
\maketitle

Because of the reliance of the particular assignments in this class on mathematical notation, 
and the fact that all assignments will be submitted electronically, 
students are encouraged to use \LaTeX{} to formalize their responses. 
\textbf{For those enrolled in the graduate section the use of latex is \emph{required}.}
This assignment (like all others) will be posted on the course \texttt{github}\footnote{\url{github.com/deblasiolab/CS4364-documents}} as source code as well as in PDF form on the course website. 
%Please submit your assignment to the professor via email, either as a link to your assignment online (i.e. overleaf or github) or as an attachment. 
Graduate students will need to include the \texttt{.tex} files as well as a PDF, this is optional but encouraged for undergraduates. 

\textbf{Question (25 points):} 
Given a segment of the genome $G$ and an RNA sequence $R$, 
describe an algorithm (along with proofs of correctness and running-time) 
to align the RNA to a segment of the genome (i.e. all of the RNA, substring of the Genome) 
allowing for gaps left by the removal of introns. 

Use the following penalty/reward structure: 
\begin{itemize}
\item Matches, mismatches, and single character indels will be scored using a function $\delta(a,b)$, defined for $a,b \in \Sigma \cup \{\texttt{-}\}$ 
(assume $\Sigma$ contains characters from both the RNA and DNA alphabets).
\item Intron (represented as long segments of deletions from the genome) will be scored not based on length but on the two characters in the RNA bordering the intron. 
Let $s_1,s_2$ and $p_1,p_2$ be the characters before and after the intron gap respectively (in the RNA of course). 
The penalty of inserting the intron between them (a substring in $R$ is $....s_1s_2p_1p_2...$) is scored by $\gamma(s_1,s_2,p_1,p_2)$.
Notice that $\gamma$ does not take into account the intron length (deletion in the genome). 
\end{itemize}

Your algorithm should have a running time of $O(m^2n)$ (assuming $|G|=m,|R|=n$) and use $O(mn)$ space.

\end{document} 

